---
title: "DATA 607 Assignment: Working with HTML and JSON"
author: "Izza Khan"
format:
  html:
    toc: true
    toc-depth: 2
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

For this warm-up exercise, I selected three books on the subject of **data science and analytics**. I manually created the same book dataset in two different file formats: an HTML table and a JSON file. I then used R to load both files into separate data frames and compared them to confirm that they contain the same information.

The three books are all relevant to this course because they focus on data wrangling, text mining, and predictive modeling.

## Data Files

This project uses two hand-built files:

- `books.html`
- `books.json`

For local knitting in the same project folder, the files can be read directly by filename. If publishing to RPubs, these files should first be uploaded to a public GitHub repository, and the URLs in the optional web-based code chunk below can be updated to the raw GitHub links.

## Load Required Packages

```{r}
library(rvest)
library(xml2)
library(jsonlite)
library(dplyr)
library(purrr)
```

## Read the HTML File into a Data Frame

```{r}
html_df <- read_html("books.html") |>
  html_element("table") |>
  html_table()

html_df
```

The HTML file stores the authors in one table cell as a semicolon-separated string, so I cleaned that column slightly to make it consistent with the JSON data after import.

```{r}
html_df <- html_df |>
  mutate(
    authors = gsub(";", ",", authors),
    authors = gsub("\\s+", " ", authors),
    publication_year = as.integer(publication_year)
  )

html_df
```

## Read the JSON File into a Data Frame

```{r}
json_raw <- fromJSON("books.json", simplifyDataFrame = TRUE)

json_df <- as_tibble(json_raw) |>
  mutate(
    authors = map_chr(authors, ~ paste(.x, collapse = ", "))
  )

json_df
```

## Compare the Two Data Frames

To compare the data frames fairly, I first made sure the column order and data types matched.

```{r}
html_df <- html_df |>
  select(title, authors, publication_year, publisher, isbn, topic)

json_df <- json_df |>
  select(title, authors, publication_year, publisher, isbn, topic)

identical_result <- identical(html_df, json_df)
all_equal_result <- all.equal(html_df, json_df)

identical_result
all_equal_result
```

## Optional: Web-Hosted Version for RPubs

When publishing to RPubs, local file paths will not work unless the source files are available on the web. After uploading `books.html` and `books.json` to a public GitHub repository, replace the placeholder URLs below with the raw GitHub links.

```{r eval=FALSE}
html_url <- "https://raw.githubusercontent.com/your-username/your-repo/main/books.html"
json_url <- "https://raw.githubusercontent.com/your-username/your-repo/main/books.json"

html_df_web <- read_html(html_url) |>
  html_element("table") |>
  html_table() |>
  mutate(
    authors = gsub(";", ",", authors),
    authors = gsub("\\s+", " ", authors),
    publication_year = as.integer(publication_year)
  )

json_df_web <- fromJSON(json_url, simplifyDataFrame = TRUE) |>
  as_tibble() |>
  mutate(
    authors = map_chr(authors, ~ paste(.x, collapse = ", "))
  )

identical(html_df_web, json_df_web)
```

## Conclusion

This exercise showed how the same dataset can be represented in both HTML and JSON formats and then loaded into R for analysis. After a small amount of cleaning—mainly standardizing the `authors` field—the two data frames matched. This demonstrates that both file types can serve as usable sources for downstream data acquisition work in R.

## Generative AI Disclosure

I used ChatGPT to help brainstorm the structure of the files, generate example code, and explain the workflow. Per the course AI policy, I would submit the full interaction transcript along with the assignment and cite the model used.

Example citation:

OpenAI. (2026). *ChatGPT (GPT-5.4 Thinking)* [Large language model]. Accessed March 11, 2026.
